Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-11 ltx/BrainIDE.tex
added appendix D for clarification

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behavior. It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology.
The model relates the average number of action potentials $g$ arriving at $r$ to the local post synaptic membrane voltage. The post-synaptic potentials generated at a neuronal population at location $r$ by action potentials arriving from all other connected populations at locations $r'$ can be described by
\begin{equation}\label{SpikesToPotential}
v\left( {r,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {r,t'} \right)dt'} + de\left( {r,t} \right).
\end{equation}
The term $de(r,t)$ can be considered the potential generated by unmodeled inputs and spontaneous background activity and is a zero mean temporal Wiener process with incremental covariance $\Sigma_e(r)dt$. The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - \alpha t)} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
g\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( v\left( r',\bar t \right) \right)dr'}
\end{equation}
where $\bar t = t - d\left| {r - r'} \right|$, $d$ is the propagation delay of action potentials, $f$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface.

@@

v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+de\left(r,t\right)

@@

\frac{dv\left( {r,t} \right)}{dt} + \alpha v\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'}  +de\left( r,t \right).

@@

\begin{eqnarray}
 \mathbf x_{1}&=&\mathbf x_{0}+T_s \mathbf Q(\mathbf x_0)\mathbf \theta-\alpha T_s\mathbf x_0+T_s\mathbf e_0 \nonumber \\
 \mathbf x_{2}&=&\mathbf x_{1}+T_s \mathbf Q(\mathbf x_1)\mathbf \theta-\alpha T_s\mathbf x_1+T_s\mathbf e_1\nonumber\\
&\vdots& \nonumber\\
 \mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1})\mathbf \theta-\alpha T_s\mathbf x_{n-1}+T_s\mathbf e_{n-1}
\end{eqnarray}
in matrix form
\begin{equation}
 \begin{bmatrix} \mathbf x_{1}\\\mathbf x_{2}\\\vdots\\\mathbf x_{n}\end{bmatrix}-\begin{bmatrix} \mathbf x_{0}\\\mathbf x_{1}\\\vdots\\\mathbf x_{n-1}\end{bmatrix}=T_s\begin{bmatrix}\mathbf Q(\mathbf x_0)&-\mathbf x_{0}\\\mathbf Q(\mathbf x_1)&-\mathbf x_{1}\\\vdots\\ \mathbf Q(\mathbf x_{n-1})&-\mathbf x_{n-1}\end{bmatrix}\begin{bmatrix}\mathbf \theta \\ \alpha\end{bmatrix}+T_s\begin{bmatrix}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{bmatrix}
\end{equation}
or in compact form we have
\begin{equation}
 \mathbf Z=\mathbf X \mathbf W+\boldsymbol \xi
\end{equation}
where
\begin{equation}
 \mathbf Z=\begin{bmatrix} \mathbf x_{1}-\mathbf x_{0}\\\mathbf x_{2}-\mathbf x_{1}\\\vdots\\\mathbf x_{n}-\mathbf x_{n-1}\end{bmatrix},\quad \mathbf X=T_s\begin{bmatrix}\mathbf Q(\mathbf x_0)&-\mathbf x_{0}\\\mathbf Q(\mathbf x_1)&-\mathbf x_{1}\\\vdots\\ \mathbf Q(\mathbf x_{n-1})&-\mathbf x_{n-1}\end{bmatrix},\quad \mathbf W=\begin{bmatrix}\mathbf \theta \\ \alpha\end{bmatrix} \text{and}\quad \boldsymbol \xi=T_s\begin{bmatrix}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{bmatrix}
\end{equation}
using the method of Least Square $ \mathbf W$ can be estimated
\begin{equation}
 \mathbf W\approx(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z
\end{equation}
% use section* for acknowledgement
\section*{Acknowledgment}
The authors would like to thank...

>>>>

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behavior. It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology.
The model relates the average number of action potentials $g$ arriving at $r$ to the local post synaptic membrane voltage. The post-synaptic potentials generated at a neuronal population at location $r$ by action potentials arriving from all other connected populations at locations $r'$ can be described by
\begin{equation}\label{SpikesToPotential}
v\left( {r,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {r,t'} \right)dt'} + e\left( {r,t} \right).
\end{equation}
The term $de(r,t)$ can be considered the potential generated by unmodeled inputs and spontaneous background activity and is a zero mean temporal Wiener process with incremental covariance $\Sigma_e(r)dt$. The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - \alpha t)} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
g\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( v\left( r',\bar t \right) \right)dr'}
\end{equation}
where $\bar t = t - d\left| {r - r'} \right|$, $d$ is the propagation delay of action potentials, $f$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface.

@@

v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)

@@

dv\left( {r,t} \right) + \alpha v\left( r,t \right)dt = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'dt}  +de\left( r,t \right).

@@

\begin{eqnarray}
 \mathbf x_{1}&=&\mathbf x_{0}+T_s \mathbf Q(\mathbf x_0)\mathbf \theta-\alpha T_s\mathbf x_0+T_s\mathbf e_0 \nonumber \\
 \mathbf x_{2}&=&\mathbf x_{1}+T_s \mathbf Q(\mathbf x_1)\mathbf \theta-\alpha T_s\mathbf x_1+T_s\mathbf e_1\nonumber\\
&\vdots& \nonumber\\
 \mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1})\mathbf \theta-\alpha T_s\mathbf x_{n-1}+T_s\mathbf e_{n-1}
\end{eqnarray}
in matrix form
\begin{equation}
 \begin{bmatrix} \mathbf x_{1}\\\mathbf x_{2}\\\vdots\\\mathbf x_{n}\end{bmatrix}-\begin{bmatrix} \mathbf x_{0}\\\mathbf x_{1}\\\vdots\\\mathbf x_{n-1}\end{bmatrix}=T_s\begin{bmatrix}\mathbf Q(\mathbf x_0)&-\mathbf x_{0}\\\mathbf Q(\mathbf x_1)&-\mathbf x_{1}\\\vdots\\ \mathbf Q(\mathbf x_{n-1})&-\mathbf x_{n-1}\end{bmatrix}\begin{bmatrix}\mathbf \theta \\ \alpha\end{bmatrix}+T_s\begin{bmatrix}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{bmatrix}
\end{equation}
or in compact form we have
\begin{equation}
 \mathbf Z=\mathbf X \mathbf W+\boldsymbol \xi
\end{equation}
where
\begin{equation}
 \mathbf Z=\begin{bmatrix} \mathbf x_{1}-\mathbf x_{0}\\\mathbf x_{2}-\mathbf x_{1}\\\vdots\\\mathbf x_{n}-\mathbf x_{n-1}\end{bmatrix},\quad \mathbf X=T_s\begin{bmatrix}\mathbf Q(\mathbf x_0)&-\mathbf x_{0}\\\mathbf Q(\mathbf x_1)&-\mathbf x_{1}\\\vdots\\ \mathbf Q(\mathbf x_{n-1})&-\mathbf x_{n-1}\end{bmatrix},\quad \mathbf W=\begin{bmatrix}\mathbf \theta \\ \alpha\end{bmatrix} \text{and}\quad \boldsymbol \xi=T_s\begin{bmatrix}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{bmatrix}
\end{equation}
using the method of Least Square $ \mathbf W$ can be estimated
\begin{equation}
 \mathbf W\approx(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z
\end{equation}
% use section* for acknowledgement
\section{From (5) to (8)}
I ignore noise here. From equation (\ref{SpikesToPotential}) we can say $h$ is the foundamental solution to the equation
\begin{equation}
 LV(r,t)=g(r,t)
\label{eq:LV}
\end{equation}
where $L$ is a temporal differential operator, since $h(t)$ is the foundamental solution to (\ref{eq:LV}) it can be written as a Green's function
\begin{equation}
 Lh(t)=\delta(t)
\end{equation}
where $L=\alpha+\frac{d}{dt}$ therefore equation (\ref{eq:LV}) becomes:
\begin{equation}
\frac{dv(r,t)}{dt}+\alpha v(r,t)=g(r,t)
\end{equation}
substituting (\ref{RateBasedInteractions}) for $g(r,t)$ gives
\begin{equation}
\frac{dv\left( {r,t} \right)}{dt} + \alpha v\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'} 
\end{equation}

\section*{Acknowledgment}
The authors would like to thank...
